Fixing Vulnerabilities Potentially Hinders Maintainability

### 2020-09-04 paper/emse20.tex
rui affiliation changed

\institute{Sofia Reis \at
			INESC ID and IST, University of Lisbon, Lisbon, Portugal\\
            \email{sofia.o.reis@tecnico.ulisboa.pt}   
			\and
			Rui Abreu \at
			INESC ID and IST, University of Lisbon, Lisbon, Portugal\\
			\email{rui@computer.org}
           \and
			Luis Cruz \at
			Delft University of Technology, Delft, The Netherlands\\
			\email{L.Cruz@tudelft.nl}
}

@@

\section{Study Implications}\label{sec:implications}
%
Our results show evidence that developers may have to reduce maintainability for the 
sake of security. We argue that developers should be able to patch and produce 
secure code without hindering the maintainability of their projects. But there are 
still concerns that need to be addressed and that this study brings awareness for:
%
\\\textit{\textbf{Tools for Patch Risk Assessment:}}
	Design debt of one guideline can lead to severe impacts on the 
	software quality~\cite{10.1145/1985362.1985366}. Some software 
	producers consider security as a first class citizen while others 
	do not. As mentioned in previous work, security it is very important 
	and should bee considered as a default feature~\cite{10.1145/2489828.2489830,kurilova2014wyvern,mcgraw2004software}. 
	However, the lack of experts or awareness to the ease of producing 
	vulnerable software leads companies to ship low quality software. 
	Providing automated tools to developers assess the risk of their 
	patches it is important to help companies shipping software of higuer quality. 
	Tools like Better Code Hub and static analysis (e.g., 
	SonarQube, Codacy, ESLint, Infer, and more) can be leveraged 
	to assess the risk of patching vulnerabilities---alongside pull
	requests/code reviews. Static code analysis may be daunting due 
	to the number of rules and different effects
	on maintainability. BCH claims to use the top-$10$ of the guidelines
	with the highest effects on maintainability. Yet, static analysis 
	can be used to complement the BCH analysis by introducing
	the capability of vulnerability detection and support the prediction 
	and prevention of the risk of a patch of hindering maintainability.
%
\\\textit{\textbf{Computer Science Curricula Needs to be Updated:}}
	Computer Science curricula is not yet prepared to properly educate 
	students for maintainable security. Students should be exposed to 
	this problem to gain experience with it. Curricula should focus on 
	the production of secure and maintainable code and alert to the 
	trade-off between both. These matters should be discussed and 
	presented to students in software engineering courses.
	Universities should also encourage students to use code quality 
	analysis tools such as BCH or similar ones. These tools have a 
	great potential to make students aware of maintainability issues 
	and beginner mistakes (e.g., coding practices violation).
%	
\\\textit{\textbf{More Secure Programming Languages Wanted:}} Programming 
	Languages should provide new design patterns to easily patch security weaknesses 
	without endangering maintainability. Ultimately, new programming languages, 
	both secure and maintainable by design, such as Wyvern~\cite{10.1145/2489828.2489830,kurilova2014wyvern}, 
	should be designed to help developers be less vulnerability prone when 
	writing and maintaining secure applications. One example is the need for 
	designing new authorization mechanisms since these are one of the most 
	complex security features to implement.
% 
% \\\textit{\textbf{Poor Coding Practices Documentation:}}
%   \Rui{too generic?}
% 	Software documentation must be provided to developers with the
% 	best practices to implement secure and maintainable software. Unmaintainable
% 	code is difficult to test, analyze and re-use. Software companies must
% 	do an effort to provide these artifacts to their developers, in order,
% 	to avoid the introduction of new vulnerabilities while patching older ones.

\section{Threats to Validity}\label{sec:threats}
%
This section presents the potential threats to validity of this study.
%
\\\textit{\textbf{Construct Validity:}} The formula was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY,Baggen2012}.
Maintainability is computed as the mean of all guidelines. Different software versions 
(vulnerable/fixed) of one vulnerability may have the same overall score
and still be affected by different guidelines.
%
\\\textit{\textbf{Internal Validity:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed while 
patching vulnerabilities. This approach discards patches that were
not explicit in commits messages. We assume that patches were performed 
using a single-commit or several sequentially. The 
perspective that a 
developer may perform a patch quickly and later 
proceed to the refactor is not considered. We assume 
that all patches were only performed once. Depending 
on the impact of the vulnerability in the system, some 
vulnerabilities may have more urgency to be fixed than 
others. Baseline commits are retrieved randomly from 
the same project of the security patch.
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style, and software contribution policies which are not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits---any kind
of commit---are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
%
\\\textit{\textbf{External Validity:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
patches in any other language but English.

@@

% \section{Plan B}
%
% Baseado nos reviewers, eu acho que para seguir as orientaçoes deles de refletir e fortalecer
% a nova assumption e as implicaçoes, teriamos de acrescentar ao Plano A, os proximos pontos
% antes de submeter à ESME.
%
% \begin{itemize}
% \item \textbf{Step 1:} Pegar nos casos em que o impacto na maintainability foi negativa e ver se foram modificados
% depois e se sim, se houve melhoria da maintainability. -> isto para então resolver a questao da
% assumption errada.
% \item \textbf{Step 2:} Se não existir nenhuma alteração, entao talvez entrar em contacto com os developers
% e perguntar se quando fizeram o patch tiveram em consideração maintainability do projecto. E se não porquê.
% \item Problemas de baselines e testes acho que nao vale a pena pegar porque vai depender
% do reviewer (eu acho).
% \end{itemize}
%
% Qual é o plano que acha melhor? Como prosseguir?
%
% A discutir, mas acho que por agora era tentar fazer o 1.
%
% \section{Later}
%
% \begin{itemize}
% 	\item \textbf{Assumptions:}\textit{authors measure the maintainability directly after the patch. inestad it should look
%   for multiple follow up commits and how maintainability evolves there, as it is likely that certain
%   issues are addresses AFTER the patching commit} -> I think we should see if later commits touch on
%   the patch and measure their maintainability again. Was a less maintainable fix pushed first and then
%   refactored, or does the code stay less maintainable?
%   \item \textbf{To back up implications:} \textit{In terms of the push for changes to curricula, I would agree that it is important
%   to always stress writing high quality code. However, the support by the study is not strong. Given a
%   majority of patches did not decrease maintainability, it shows that there is a reasonable focus on maintainability.
%   Similarly, it is important to understand the circumstances when it decreased. As previously mentioned, organizations
%   like would prioritize getting the system patched faster (especially in the case of a severe vulnerability) over
%   its maintainability.} -> Work with practioners to understand if they were taking into consideration maintainability?
%   \item \textbf{Baseline:}\textit{Authors just search for commits of a similar size. There are other important aspects
%   that are entirely neglected such as size of the changed method or how complex it it (by means of
%   metrics)} -> Improve baseline.
%   \item \textbf{Baseline:}\textit{Similarly, the comparison to “regular” commits really compares security patches to
%   any other type of change. It would have been more meaningful to compared them to specific types of
%   commits, like bug fixes (e.g., are security fixes less maintainable than bug fixes?).}
%   \item \textbf{Tests:}\textit{or the chosen statistical test, the hypothesis may be not well chosen as it just checks for
%   the same distribution of pairs of versions. However, this means if a similar number of pairs have
%   a negative and a positive impact, the hypothesis holds however, if this is not the case, the statistical
%   test does not state in which direction the hypothesis is rejected, that is, what eventually is the
%   alternative hypothesis. But this is necessary to get insights whether maintainability is hampered or not.
%   Authors may rethink the choice of the hypothesis or even the test itself.} -> What about this?
%   \item \textbf{Thesis idea (PL):} It is important
%   to understand the correlation between the severity of the vulnerability and the programming language as well
%   as the maintainability impact. It is not clear how software written a more security-focused language would
%   inherently be more maintainable (it assumes the flaw is the language and not the person).
% \end{itemize}

>>>>

\institute{Sofia Reis \at
			INESC ID and IST, University of Lisbon, Lisbon, Portugal\\
            \email{sofia.o.reis@tecnico.ulisboa.pt}   
			\and
			Rui Abreu \at
			INESC ID and FEUP, University of Porto, Porto, Portugal\\
			\email{rui@computer.org}
           \and
			Luis Cruz \at
			Delft University of Technology, Delft, The Netherlands\\
			\email{L.Cruz@tudelft.nl}
}

@@

\section{Study Implications}\label{sec:implications}
%
Our results show evidence that developers may have to reduce maintainability for the 
sake of security. We argue that developers should be able to patch and produce 
secure code without hindering the maintainability of their projects. But there are 
still concerns that need to be addressed and that this study brings awareness for:
%

\textit{\textbf{Tools for Patch Risk Assessment:}}
	Design debt of one guideline can lead to severe impacts on the 
	software quality~\cite{10.1145/1985362.1985366}. Some software 
	producers consider security as a first class citizen while others 
	do not. As mentioned in previous work, security it is very important 
	and should bee considered as a default feature~\cite{10.1145/2489828.2489830,kurilova2014wyvern,mcgraw2004software}. 
	However, the lack of experts or awareness to the ease of producing 
	vulnerable software leads companies to ship low quality software. 
	Providing automated tools to developers assess the risk of their 
	patches it is important to help companies shipping software of higuer quality. 
	Tools like Better Code Hub and static analysis (e.g., 
	SonarQube, Codacy, ESLint, Infer, and more) can be leveraged 
	to assess the risk of patching vulnerabilities---alongside pull
	requests/code reviews. Static code analysis may be daunting due 
	to the number of rules and different effects
	on maintainability. BCH claims to use the top-$10$ of the guidelines
	with the highest effects on maintainability. Yet, static analysis 
	can be used to complement the BCH analysis by introducing
	the capability of vulnerability detection and support the prediction 
	and prevention of the risk of a patch of hindering maintainability.
%

\textit{\textbf{Computer Science Curricula Needs to be Updated:}}
	Computer Science curricula is not yet prepared to properly educate 
	students for maintainable security. Students should be exposed to 
	this problem to gain experience with it. Curricula should focus on 
	the production of secure and maintainable code and alert to the 
	trade-off between both. These matters should be discussed and 
	presented to students in software engineering courses.
	Universities should also encourage students to use code quality 
	analysis tools such as BCH or similar ones. These tools have a 
	great potential to make students aware of maintainability issues 
	and beginner mistakes (e.g., coding practices violation).
%

\textit{\textbf{More Secure Programming Languages Wanted:}} Programming 
	Languages should provide new design patterns to easily patch security weaknesses 
	without endangering maintainability. Ultimately, new programming languages, 
	both secure and maintainable by design, such as Wyvern~\cite{10.1145/2489828.2489830,kurilova2014wyvern}, 
	should be designed to help developers be less vulnerability prone when 
	writing and maintaining secure applications. One example is the need for 
	designing new authorization mechanisms since these are one of the most 
	complex security features to implement.
% 
% \\\textit{\textbf{Poor Coding Practices Documentation:}}
%   \Rui{too generic?}
% 	Software documentation must be provided to developers with the
% 	best practices to implement secure and maintainable software. Unmaintainable
% 	code is difficult to test, analyze and re-use. Software companies must
% 	do an effort to provide these artifacts to their developers, in order,
% 	to avoid the introduction of new vulnerabilities while patching older ones.

\section{Threats to Validity}\label{sec:threats}
%
This section presents the potential threats to validity of this study.
%

\textit{\textbf{Construct Validity:}} The formula was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY,Baggen2012}.
Maintainability is computed as the mean of all guidelines. Different software versions 
(vulnerable/fixed) of one vulnerability may have the same overall score
and still be affected by different guidelines.
%

\textit{\textbf{Internal Validity:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed while 
patching vulnerabilities. This approach discards patches that were
not explicit in commits messages. We assume that patches were performed 
using a single-commit or several sequentially. The 
perspective that a 
developer may perform a patch quickly and later 
proceed to the refactor is not considered. We assume 
that all patches were only performed once. Depending 
on the impact of the vulnerability in the system, some 
vulnerabilities may have more urgency to be fixed than 
others. Baseline commits are retrieved randomly from 
the same project of the security patch.
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style, and software contribution policies which are not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits---any kind
of commit---are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
%

\textit{\textbf{External Validity:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
patches in any other language but English.

@@

% \section{Plan B}
%
% Baseado nos reviewers, eu acho que para seguir as orientaçoes deles de refletir e fortalecer
% a nova assumption e as implicaçoes, teriamos de acrescentar ao Plano A, os proximos pontos
% antes de submeter à ESME.
%
% \begin{itemize}
% \item \textbf{Step 1:} Pegar nos casos em que o impacto na maintainability foi negativa e ver se foram modificados
% depois e se sim, se houve melhoria da maintainability. -> isto para então resolver a questao da
% assumption errada.
% \item \textbf{Step 2:} Se não existir nenhuma alteração, entao talvez entrar em contacto com os developers
% e perguntar se quando fizeram o patch tiveram em consideração maintainability do projecto. E se não porquê.
% \item Problemas de baselines e testes acho que nao vale a pena pegar porque vai depender
% do reviewer (eu acho).
% \end{itemize}
%
% Qual é o plano que acha melhor? Como prosseguir?
%
% A discutir, mas acho que por agora era tentar fazer o 1.
%
% \section{Later}
%
% \begin{itemize}
% 	\item \textbf{Assumptions:}\textit{authors measure the maintainability directly after the patch. inestad it should look
%   for multiple follow up commits and how maintainability evolves there, as it is likely that certain
%   issues are addresses AFTER the patching commit} -> I think we should see if later commits touch on
%   the patch and measure their maintainability again. Was a less maintainable fix pushed first and then
%   refactored, or does the code stay less maintainable?
%   \item \textbf{To back up implications:} \textit{In terms of the push for changes to curricula, I would agree that it is important
%   to always stress writing high quality code. However, the support by the study is not strong. Given a
%   majority of patches did not decrease maintainability, it shows that there is a reasonable focus on maintainability.
%   Similarly, it is important to understand the circumstances when it decreased. As previously mentioned, organizations
%   like would prioritize getting the system patched faster (especially in the case of a severe vulnerability) over
%   its maintainability.} -> Work with practioners to understand if they were taking into consideration maintainability?
%   \item \textbf{Baseline:}\textit{Authors just search for commits of a similar size. There are other important aspects
%   that are entirely neglected such as size of the changed method or how complex it it (by means of
%   metrics)} -> Improve baseline.
%   \item \textbf{Baseline:}\textit{Similarly, the comparison to “regular” commits really compares security patches to
%   any other type of change. It would have been more meaningful to compared them to specific types of
%   commits, like bug fixes (e.g., are security fixes less maintainable than bug fixes?).}
%   \item \textbf{Tests:}\textit{or the chosen statistical test, the hypothesis may be not well chosen as it just checks for
%   the same distribution of pairs of versions. However, this means if a similar number of pairs have
%   a negative and a positive impact, the hypothesis holds however, if this is not the case, the statistical
%   test does not state in which direction the hypothesis is rejected, that is, what eventually is the
%   alternative hypothesis. But this is necessary to get insights whether maintainability is hampered or not.
%   Authors may rethink the choice of the hypothesis or even the test itself.} -> What about this?
%   \item \textbf{Thesis idea (PL):} It is important
%   to understand the correlation between the severity of the vulnerability and the programming language as well
%   as the maintainability impact. It is not clear how software written a more security-focused language would
%   inherently be more maintainable (it assumes the flaw is the language and not the person).
% \end{itemize}
\pagebreak
